---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r load}
setwd("~/Development/github/RepData_PeerAssessment1")
activity <- read.csv("activity.csv")
str(activity)
summary(activity)
```

## Process the data

```{r process}
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?

Total number of steps each day

```{r stepsPerDay}
stepsPerDay <- aggregate(activity$steps, list(activity$date), sum, na.rm=TRUE)
names(stepsPerDay) <- c("date", "steps")
library(ggplot2)
ggplot(stepsPerDay, aes(x=steps)) + geom_histogram(binwidth=5000)
meanSteps <- as.integer(mean(stepsPerDay$steps))
meanSteps
medianSteps <- median(stepsPerDay$steps)
medianSteps
```

## What is the average daily activity pattern?

```{r stepsDailyPattern}
stepsPattern <- aggregate(activity$steps, list(activity$interval), mean, na.rm=TRUE)
names(stepsPattern) <- c("time", "steps")
plot(stepsPattern, type="l", main="Average daily activity pattern")
maxInterval = stepsPattern$time[which.max(stepsPattern$steps)]
```

The 5-minute interval, with on average across all days, maximum number of steps: `r maxInterval`. 

## Imputing missing values

```{r NAs}
NAs <- sum(is.na(activity))
rows <- nrow(activity)
NAratio <- NAs/rows
```

There are `r NAs` missing values out of a total of `r rows`, or `r NAratio*100` %.

Strategy: fill in missing values with the mean for that 5-minute interval.

```{r impute}
means <- aggregate(steps ~ interval, data = activity, mean, na.rm=TRUE)
names(means)[2] <- "impute.steps"
activityFix <- merge(activity, means) 
activityFix$steps[is.na(activityFix$steps)] <- activityFix$impute.steps[is.na(activityFix$steps)]
activityFix$impute.steps <- NULL
str(activityFix)
summary(activityFix)
activity$na = "missing"
activityFix$na = "imputed"
combo = rbind(activity, activityFix)
library(ggplot2)
comboPerDay = aggregate(combo$steps, list(combo$date, combo$na), sum, na.rm=TRUE)
colnames(comboPerDay) = c("date", "na", "steps")
ggplot(comboPerDay, aes(x=steps, fill=na)) + geom_histogram(position="identity", alpha=.5, binwidth=5000)
```

8 days had missing values for all of the intervals, which are now filled with the mean values per interval,

The mean and median shift accordingly:

```{r}
meanSteps
meanSteps2 <- as.integer(mean(comboPerDay$steps[comboPerDay$na == "imputed"]))
meanSteps2
medianSteps
medianSteps2 <- median(comboPerDay$steps[comboPerDay$na == "imputed"])
medianSteps2
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activityFix$date = as.Date(activityFix$date)
activityFix$weekday = ifelse(weekdays(activityFix$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
stepsPattern2 <- aggregate(activityFix$steps, list(activityFix$interval, activityFix$weekday), mean, na.rm=TRUE)
names(stepsPattern2) = c("interval", "weekday", "steps")
ggplot(stepsPattern2, aes(x=interval, y=steps)) + geom_line() + facet_grid(weekday ~.)
```
